---
title: "individual2.Rmd"
author: "Mike Brown"
date: "4/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse) 
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(classdata)
library(plotly)
library(ggrepel)

acc <- read.csv("C:/Users/brown/OneDrive/Desktop/DS202/Lab 5/ds202_lab5/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("C:/Users/brown/OneDrive/Desktop/DS202/Lab 5/ds202_lab5/person.csv", stringsAsFactors = FALSE)
```
1.Are there some days of the week where more accidents happen than the others (see FARS manual, use variable `DAY_WEEK`)?

Answer: 
```{r}
acc1 <- wday(acc$DAY_WEEK, label = TRUE)
as.data.frame(table(acc1))
```
Saturday has more accidents than any other day of the week.

2.Create a data frame containing the persons who are fatally hurt in the accidents (look up variable `INJ_SEV`).

Answer:
```{r}
fatal <- filter(per, INJ_SEV == 4)
head(fatal)
```
3.Create a data frame containing the most dangerous vechitle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable `MAKE`)
```{r}
GLC <- readxl::read_xlsx("C:/Users/brown/OneDrive/Desktop/DS202/Lab 5/ds202_lab5/FRPP_GLC_-_United_StATESaPRIL62021.xlsx")
```
Answer:


4.Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the *most dangerous vechicle*, and what you find from the map. (Hint: Read the description for the `STATE` and `COUNTY` columns in the FARS manual. The state and &county codes are Geographic Locator Codes (GLCs) from the General Services Administration's (GSA) publication. Use `readxl::read_xlsx` to read in the GLCs.)

Answer:
```{r}
perMDV <- fatal %>% select(STATE, MAKE, INJ_SEV) %>% na.omit() %>% group_by(STATE, MAKE) %>% summarise(count = n())

perMDVRes <- perMDV %>% group_by(STATE) %>% top_n(1, count)
accLongLatAVG <- acc %>%
  group_by(STATE) %>%
  summarise(avgLong = mean(LONGITUD), avgLat = mean(LATITUDE))

perMDVResLatLong <- full_join(perMDVRes, accLongLatAVG, by = "STATE")
```
5.Join the accident and person table (work out which varaible(s) to use)

Answer:
```{r}
join <-left_join(per, acc, by="ST_CASE")
head(join)
```
6.Tally the number of accidents by day of the week (Day_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results and explain what you find.

Answer:
```{r}
join$DAY_WEEK <- wday(join$DAY_WEEK, label=TRUE)
tally<- summarize(group_by(join,DAY_WEEK, HOUR.x, SEX),count=n())
ggplot(tally, aes(x=HOUR.x, y=count, group=SEX))+geom_line(color=tally$SEX)+geom_text_repel(aes(x=HOUR.x,y=count,label=SEX))+facet_wrap(~DAY_WEEK)+xlim(0,24)
```
This faceted graph shows that males are more likely to be in an accident no matter what the time is. This also shows that crashes are more frequent on the weekends closer to the end of the day. This is likely due to people partying on their days off and attempting to drive home when they're done.

7.Now plot a choropleth map of the number of deaths on a county level. Also explain what you find.


Answer:



8. Is summer or winter more dangerous? Does this depend on states? Explore and explain.



Answer:
