---
title: "individual2.Rmd"
author: "Mike Brown"
date: "4/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse) 
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(classdata)
library(plotly)
library(ggrepel)

acc <- read.csv("C:/Users/brown/OneDrive/Desktop/DS202/Lab 5/ds202_lab5/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("C:/Users/brown/OneDrive/Desktop/DS202/Lab 5/ds202_lab5/person.csv", stringsAsFactors = FALSE)
```
1.Are there some days of the week where more accidents happen than the others (see FARS manual, use variable `DAY_WEEK`)?

Answer: 
```{r}
acc1 <- wday(acc$DAY_WEEK, label = TRUE)
as.data.frame(table(acc1))
```
Saturday has more accidents than any other day of the week.

2.Create a data frame containing the persons who are fatally hurt in the accidents (look up variable `INJ_SEV`).

Answer:
```{r}
fatal <- filter(per, INJ_SEV == 4)
head(fatal)
```
3.Create a data frame containing the most dangerous vechitle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable `MAKE`)


Answer:


4.Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the *most dangerous vechicle*, and what you find from the map. (Hint: Read the description for the `STATE` and `COUNTY` columns in the FARS manual. The state and &county codes are Geographic Locator Codes (GLCs) from the General Services Administration's (GSA) publication. Use `readxl::read_xlsx` to read in the GLCs.)

Answer:


5.Join the accident and person table (work out which varaible(s) to use)


Answer:


6.Tally the number of accidents by day of the week (Day_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results and explain what you find.


Answer:


7.Now plot a choropleth map of the number of deaths on a county level. Also explain what you find.


Answer:



8. Is summer or winter more dangerous? Does this depend on states? Explore and explain.



Answer:
