---
title: "team.Rmd"
author: "Mike Brown and Nathan Marquardt"
date: "4/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.Are there some days of the week where more accidents happen than the others (see FARS manual, use variable `DAY_WEEK`)?


Answer: 
```{r}
library(tidyverse) 
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(classdata)
library(plotly)
library(ggrepel)

acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

acc$DAY_WEEK <- wday(acc$DAY_WEEK, label = TRUE)

accCount <- acc %>%
            group_by(DAY_WEEK) %>%
            summarise(count = n())

ggplot(aes(x = DAY_WEEK, weight = count, fill = DAY_WEEK), data = accCount) + geom_bar() + xlab("Day of the Week") + ylab("Number of accidents") + ggtitle("Number of accidents per weekday")
```
Based off of the graph above we can see that Friday - Sunday are the days with the most accidents.

2.Create a data frame containing the persons who are fatally hurt in the accidents (look up variable `INJ_SEV`).


Answer:
```{r}
fatal <- filter(per, INJ_SEV == 4)
head(fatal)
```
3.Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable `MAKE`)


Answer:
```{r}
perMDV <- fatal %>%
          select(STATE, MAKE, INJ_SEV) %>%
          na.omit() %>%
          group_by(STATE, MAKE) %>%
          summarise(count = n())
         


perMDVRes <- perMDV %>%
              group_by(STATE) %>%
              top_n(1, count)
head(perMDVRes)
```

4.Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the *most dangerous vechicle*, and what you find from the map. (Hint: Read the description for the `STATE` and `COUNTY` columns in the FARS manual. The state and &county codes are Geographic Locator Codes (GLCs) from the General Services Administration's (GSA) publication. Use `readxl::read_xlsx` to read in the GLCs.)


Answer:
```{r}
s <- map_data("state")
u <- fatal %>% group_by(STATE,MAKE) %>% right_join(s,by=c('STATE'='group')) %>%              summarize(avgLong=mean(long),avgLat=mean(lat))
ggplot(s,aes(x='long',y='lat'))+geom_polygon(aes(group=group))+geom_text(aes(label=MAKE),data=u)
```
Following the posts in piazza, line for line, we can see how to get the answer but for some reason this is all we can get. 

5.Join the accident and person table (work out which varaible(s) to use)


Answer:
```{r}
join <-left_join(per, acc, by="ST_CASE")
head(join)
```

6.Tally the number of accidents by day of the week (Day_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results and explain what you find.


Answer:
```{r}
tally2<-join %>% group_by(DAY_WEEK, HOUR.x, SEX) %>% summarize(count=n())
tally<- summarize(group_by(join,DAY_WEEK, HOUR.x, SEX),count=n())
ggplot(tally, aes(x=HOUR.x, y=count, group=SEX))+geom_line(color=tally$SEX)+geom_text_repel(aes(x=HOUR.x,y=count,label=SEX))+facet_wrap(~DAY_WEEK)+xlim(0,24)
```
This faceted graph shows that males (1) are more likely to be in an accident no matter what the time is. This also shows that crashes are more frequent on the weekends closer to the end of the day. This is likely due to people partying on their days off and attempting to drive home when they're done.

7.Now plot a choropleth map of the number of deaths on a county level. Also explain what you find.


Answer:
```{r}
s <- map_data("state")
c <- map_data("county")
g <- fatal %>% group_by(COUNTY) %>% summarize(count=n())
cg <- full_join(c,g,by=c('group'='COUNTY'))
ggplot(cg,aes(x=long,y=lat))+geom_polygon(aes(group=group, fill = count))
```
8. Is summer or winter more dangerous? Does this depend on states? Explore and explain.


Answer:
```{r}
accWin <- acc %>%
          filter(MONTH == 12 | MONTH == 1 | MONTH == 2)

accSum <- acc %>%
          filter(MONTH == 6 | MONTH == 7 | MONTH == 8 | MONTH == 9)

length(accWin$ST_CASE)
length(accSum$ST_CASE)

perWin <- per %>%
                filter(MONTH == 12 | MONTH == 1 | MONTH == 2)

perSum <- per %>%
            filter(MONTH == 6 | MONTH == 7 | MONTH == 8 | MONTH == 9)

perWinFatal <- perWin %>%
                  filter(INJ_SEV == 4)

perSumFatal <- perSum %>%
                filter(INJ_SEV == 4)

percWinFatal <- length(perWinFatal$MONTH) / length(perWin$MONTH)
percSumFatal <- length(perSumFatal$MONTH) / length(perSum$MONTH)

perWinCount <- perWin %>%
                group_by(STATE, INJ_SEV) %>%
                summarise(win_Count = n())

perSumCount <- perSum %>%
                group_by(STATE, INJ_SEV) %>%
                summarise(sum_Count = n())

perInjCount <- full_join(perWinCount, perSumCount, by = c("STATE", "INJ_SEV"))

perWinStateCount <- perWin %>%
                      group_by(STATE) %>%
                      summarise(winST_count = n())

perSumStateCount <- perSum %>%
                    group_by(STATE) %>%
                    summarise(sumST_count = n())

perWinSumSTCount <- full_join(perWinStateCount, perSumStateCount, by = "STATE")

perInjCountPer <- full_join(perInjCount, perWinSumSTCount, by = "STATE")

perInj <- perInjCountPer %>%
          mutate(winPercentage = win_Count / winST_count, sumPercentage = sum_Count / sumST_count)

perfatal <- perInj %>% filter(INJ_SEV == 4)
perfatal
```

From the tables above, we noted that there were more accidents in the summer (Jun-Sep) versus the winter(Dec-Feb). However, we also noted that the fatality percentage overall between winter and summer for all states grouped together were nearly identical. It seems like states with typically warmer weather have higher fatality percentages in the winter and typically colder states have higher fatality percentages in the summer. For the pattern we see with the warmer states, this could be due to them having more difficulties driving in winter conditions. For the colder states, they receive an increase in tourism during the summer which will inherently increase the chances of a wreck.